
# Simulation Study

## Required Packages and Help Functions

```{r}
#| label: packages
#| eval: true
#| echo: true
#| code-fold: true

# Load required packages
suppressPackageStartupMessages({
  # General
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(parallel)
  library(furrr)
  library(tibble)
  library(Rcpp)
  library(DT)
  library(ggplot2)
  library(clustermq)
  library(foreach)
  library(doRNG)
  library(kableExtra)
  # Specific Analyses
  library(RobinCar)
  library(Exact)
  library(sandwich)
  library(margins)
  library(marginaleffects)
  library(detectseparation)
  library(logistf)
  #library(fastglm)
})

# Source helper scripts

# source helper scripts from R/
for (f in list.files("R", pattern = "*.R", full.names = TRUE)) {
  source(f)
}

# source helper scripts from src/
for (f in list.files("src", pattern = "*.cpp", full.names = TRUE)) {
  sourceCpp(f)
}

# Set cluster options
options(clustermq.scheduler = "slurm")
```

## Simulation Parameters

We specify the following parameters to define the different simulation scenarios:

```{r}
#| label: simScenarios
#| eval: true
#| echo: true

# Total sample sizes
n <- seq(30, 150, 30)

# Proportion/prevalence/response rate in control group
prev <- .2

# Difference of treatment group to control
diff <- c(0, .15, .3)

# Define distribution for each covariate
cov_rng <- list(
  c("rbinom", "rbinom")
)

# Set parameters for distribution functions specified above (skip n)
cov_pars <- list(
  list(
    c(size = 1, prob = .5),
    c(size = 1, prob = .5)
  )
)

# True regression weights of covariates
cov_coefs <- list(
  "none" = log(c(1, 1)),
  "medium" = log(c(1.5, 1.5)),
  "high" = log(c(3, 3))
)

# Randomization scheme
rand <- "simple"

# Treatment allocation
treat_alloc <- list(
  c(1,1)
)

# Create all possible factor combinations
scenarios <- expand_grid(
  n,
  prev,
  diff,
  cov_rng,
  cov_pars,
  cov_coefs,
  rand, 
  treat_alloc
)
scenarios <- scenarios |> 
  mutate(
    pop_coefs = purrr::pmap(scenarios[,2:6], def_population_coefs)
  )

rm(n, diff, prev, cov_rng, cov_pars, cov_coefs, rand, treat_alloc)
```

To evaluate the chosen parameter values, we can check the implied (marginal) probabilities across treatment groups and covariates:

```{r}
#| label: tabulateProbs
#| eval: true
#| echo: true
#| code-fold: true

# define helper function
get_prob <- function(x1, x2, z, pop_coefs, cov_coefs) {
  
  b0 <- unlist(pop_coefs)[1]
  bz <- unlist(pop_coefs)[2]
  b1 <- unlist(cov_coefs)[1]
  b2 <- unlist(cov_coefs)[2]
  plogis(b0 + z * bz + x1 * b1 + x2 * b2)
  
}

# output matrix
out <- matrix(nrow = 18, ncol = 7)

for (s in 1:9) {
  
  # temporary array of probabilities
  tmp <- array(dim = c(2, 2, 2))
  
  for (i in 1:2) {
    tmp[,,i] <- outer(
      X = c(0, 1),
      Y = c(0, 1),
      FUN = get_prob,
      z = c(0, 1)[i],
      pop_coefs = scenarios[s, "pop_coefs"],
      cov_coefs = scenarios[s, "cov_coefs"]
    )
  }
  
  # calculate marginal probabilities and marginal effect
  marg_x_0 <- mean(tmp[1,,]) |> rep(2)
  marg_x_1 <- mean(tmp[2,,]) |> rep(2)
  me_x <- (mean(tmp[2,,]) - mean(tmp[1,,])) |> rep(2)
  
  # combine
  out[((s - 1) * 2 + 1):(s*2),] <- cbind(
    tmp[,,1], tmp[,,2], 
    marg_x_0, marg_x_1, me_x
  ) |> round(digits = 2)
  
}

# Prepare output table 
data.frame(
  rep(c("0", "0.15", "0.30"), each = 6),
  rep(c("log(1)", "log(1.5)", "log(3)"), each = 2, times = 3),
  rep(c("x1 = 0", "x1 = 1"), times = 9),
  out
) |> 
  kable(
    "html",
    col.names = c("\\[ \\text{RD}_z \\]", "\\[ \\beta_x \\]", "\\[ x_1\\]", rep(c("\\[ x_2 = 0\\]", "\\[ x_2 = 1\\]"), 2), 
                  "\\[ \\bar{x = 0} \\]", "\\[ \\bar{x = 1} \\]", "\\[ \\text{RD}_x \\]"),
    escape = TRUE,
    digits = 2,
    align = "c"
  ) |>
  kableExtra::collapse_rows(columns = c(1, 2, 8, 9, 10), valign = "middle") |> 
  kableExtra::add_header_above(c(" " = 3, "z = 0" = 2, "z = 1" = 2, " " = 3)) |>
  kableExtra::kable_styling()

```



## Run the Simulation

```{r}
#| label: runSimulation
#| eval: true
#| echo: true
#| cache: true
#| cache-lazy: false
#| message: false
#| warning: false

# 1) Use an RNG suitable for parallel streams
RNGkind("L'Ecuyer-CMRG")   # Recommended for reproducible parallel RNG

# 2) Set a single, top-level seed ONCE
set.seed(20260119)

# 3) Set simulation parameters
chunk_size <- 1250
replications <- 5e4
chunks <- ceiling(replications / chunk_size)
total_jobs <- nrow(scenarios) * chunks

# 4) Register clustermq backend
register_dopar_cmq(
  n_jobs = total_jobs,
  memory = 4096,
  export = mget(ls(envir = .GlobalEnv), envir = .GlobalEnv)
)

# 5) Start simulation
(start_time <- Sys.time())

res <- foreach(job = 1:total_jobs, .combine = 'rbind',
               .options.RNG = 2026011) %dorng% {
  
  # Compute scenario and chunk index
  scenario_id <- ((job - 1) %/% chunks) + 1
  chunk_id <- ((job - 1) %% chunks) + 1
  
  # Run the chunk
  single_scenario(
    n_sim = chunk_size,
    params = scenarios[scenario_id, ],
    methods = c(
      "compute_suissa_shuster", "compute_cmh", "compute_mh_sato",
      "compute_mh_mgr", "compute_ge", "compute_ye", "compute_liu",
      "compute_g_firth", "compute_score_test", "compute_boot"
    ),
    alpha = 0.05,
    type_Ge = "orig",
    type_Liu = "HC3",
    flic = TRUE,
    R = 1000,
    boot_strat = FALSE
  )
  
}

# End simulation
(end_time <- Sys.time())

# Total duration
end_time - start_time

```

Let's save the results:

```{r}
#| label: saveResults
#| eval: true
#| echo: true
#| cache: true
#| code-fold: true

saveRDS(
  object = res,
  file = "sim_results.rds"
)
```


```{r}
#| label: temp
#| echo: false
#| eval: false

res <- readRDS("sim_results.rds")
```


## Results

```{r}
#| label: results
#| eval: true
#| echo: true
#| cache: true
#| cache-lazy: false
#| code-fold: true

res <- res |> 
  mutate(
    id = rep(seq_len(replications * nrow(scenarios)), each = 10),
    cov_prog_val = sapply(
      `list(cov_coefs)`, 
      \(x) ifelse(
        identical(x, log(rep(1, 2))), 
        "none",
        ifelse(
          identical(x, log(rep(1.5, 2))), 
          "medium",
          "high"
        )
      )
    ),
    bias = rd_est - diff,
    coverage = dplyr::between(diff, rd_lcl, rd_ucl),
    reject = rd_pval < .05
  ) |> 
  group_by(id) |> 
  mutate(
    converged = na.omit(converged) |> unique(),
    separation = na.omit(separation) |> unique(),
    separation_treatment = na.omit(separation_treatment) |> unique()
  )
```


### Convergence Issues/Separation

Full table of results:

```{r}
#| label: checkIssues
#| eval: true
#| echo: true
#| code-fold: true

res |> 
  group_by(n, diff, cov_prog_val) |> 
  summarize(
    convergence = mean(!converged),
    separation = mean(separation),
    .groups = "drop"
  ) |> 
  datatable(options = list(
    scrollX = TRUE,
    scrollY = "400px",
    pageLength = 24
  ))
```

Visualization:

```{r}
#| label: plotConvergence
#| echo: false
#| eval: true

# res |> 
#   group_by(n, diff, cov_prog_val) |> 
#   summarize(
#     separation = mean(separation),
#     .groups = "drop"
#   ) |> 
#   filter(n < 100) |> 
#   mutate(rd = case_when(
#     diff == 0 ~ "Risk Difference = 0",
#     diff == 0.15 ~ "Risk Difference = 0.15",
#     TRUE ~ "Risk Difference = 0.30"
#   )) |> 
#   ggplot(aes(x = n, y = separation, linetype = cov_prog_val, color = cov_prog_val)) +
#   facet_wrap(~ rd) +
#   geom_line(lwd = .75) +
#   geom_point(size = 2) +
#   scale_x_continuous(
#     limits = c(30, 90),
#     breaks = c(30, 60, 90)
#   ) +
#   scale_linetype_discrete(
#     "Prog. Val. Covariates"
#   ) +
#   scale_color_discrete(
#     "Prog. Val. Covariates"
#   ) +
#   labs(
#     x = "Total Sample Size",
#     y = "% Convergence Issues"
#   ) +
#   theme_bw() +
#   theme(
#     text = element_text(size = 10, color = "black"),
#     legend.position = "bottom",
#     legend.key.width = unit(1.5, "lines"),
#     axis.text = element_text(size = 10, color = "black")
#   )

rd <- c(
  "0"    = "delta==0",
  "0.15" = "delta==0.15",
  "0.3" = "delta==0.30"
)

labeller_rd <- as_labeller(rd, label_parsed)

res |>
  group_by(n, diff, cov_prog_val) |>
  summarize(
    separation = mean(separation),
    .groups = "drop"
  ) |>
  mutate(diff = as.character(diff)) |>
  filter(n < 100) |>
  mutate(cov_prog_val = factor(
    cov_prog_val, 
    levels = c("high", "medium", "none"),
    labels = c("log(3)", "log(1.5)", "log(1)"))) |> 
  ggplot(aes(x = n, y = separation,
             linetype = cov_prog_val,
             shape = cov_prog_val,
             color = cov_prog_val)) +
  facet_wrap(~ diff, labeller = labeller_rd) +
  geom_line(lwd = .75) +
  geom_point(size = 3) +
  scale_x_continuous(
    limits = c(30, 90),
    breaks = c(30, 60, 90)
  ) +
  scale_linetype_manual(
    expression(paste(beta[1],",",beta[2])),
    values = c("solid", "dashed", "dotted")
  ) +
  scale_shape_manual(
    expression(paste(beta[1],",",beta[2])),
    values = c(15, 17, 16)
  ) +
  scale_color_manual(
    expression(paste(beta[1],",",beta[2])),
    values = c("firebrick", "red", "darksalmon")
  ) +
  labs(
    x = "Total Sample Size",
    y = "% Separation Issues"
  ) +
  theme_bw() +
  theme(
    text = element_text(size = 12, color = "black"),
    legend.position = "bottom",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 12, color = "black")
  )

```

### Power, Bias, & Coverage

```{r}
#| label: checkResults
#| eval: true
#| echo: true
#| code-fold: true

res_agg <- res |> 
  mutate(sep_cat = ifelse(separation, ifelse(separation_treatment, "treatment", "covariate"), "none")) |> 
  filter(converged) |> 
  group_by(n, diff, cov_prog_val, sep_cat, method) |>
  summarize(
    bias_mean = mean(bias, na.rm = TRUE),
    rmse = sqrt(mean(bias^2, na.rm = TRUE)),
    coverage = mean(coverage, na.rm = TRUE),
    power = mean(reject, na.rm = TRUE),
    n_cases = n(),
    .groups = "drop"
  )

# res_agg |>
#   datatable(options = list(
#     scrollX = TRUE,
#     scrollY = "400px",
#     pageLength = 100
#   )) |>
#   formatRound(columns = 6:9, digits = 3)

```


### Figures: Type I Error/Power

```{r}
#| label: preparePlotData
#| echo: false
#| eval: true
#| cache: true

plot_dat <- res |> 
  mutate(cov_prog_val = factor(cov_prog_val, 
                                    levels = c("none", "medium", "high"),
                                    labels = c("Prog. Val: None",
                                               "Prog. Val: Medium",
                                               "Prog. Val: High"))
  ) |> 
  filter(converged) |> 
  group_by(n, diff, cov_prog_val, method) |>
  summarize(
    bias_mean = mean(bias, na.rm = TRUE),
    rmse = sqrt(mean(bias^2, na.rm = TRUE)),
    coverage = mean(coverage, na.rm = TRUE),
    power = mean(reject, na.rm = TRUE),
    n_cases = n(),
    .groups = "drop"
  )

```

```{r}
#| label: plotOptions
#| echo: false
#| eval: true

# Gather methods and organize in clusters
breaks <- c(
  "Ge (orig)", "Liu (HC3)", "Ye", "Firth (FLIC)", "g-comp bootstrap", "Score Test",
  "MH (mGR)", "MH (Sato)", "cmh test", "suissa-shuster"
)

# Set labels
labels <- c(
  "G-Comp.: Ge",
  "G-Comp.: Liu",
  "G-Comp.: Ye",
  "G-Comp.: Firth",
  "G-Comp.: Bootstrap",
  "G-Comp.: Score Test",
  "Mantel-Haenszel RD: modified GR",
  "Mantel-Haenszel RD: Sato",
  "Cochran-Mantel-Haenszel Test",
  "Suissa-Shuster Test"
)

# Set colors: unique per method, grouped by clusters
values <- c(
  "dodgerblue4", "lightskyblue1", "deepskyblue", "dodgerblue", "#5DADE2", "blue",  # G-Comp (blue family)
  "#E69F00", "#F0E442",                                              # MH (orange/yellow)
  "#CC79A7",                                                         # Suissa-Shuster (magenta)
  "#009E73"                                                          # CMH (green)
)

# Linetypes: Same within cluster, different across clusters
line_types <- c(
  rep("solid", 6),      # G-Comp
  rep("dashed", 2),     # MH
  "dotdash",            # CMH
  "dotted"              # Suissa-Shuster
)

# Shapes: Unique within clusters
shapes <- c(19, 17, 15, 3, 8, 18, 19, 17, 15, 19)
```


#### Type I Error (True Null Effect)

```{r}
#| label: plotOverallPowerNull
#| echo: false
#| eval: true

plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = power, group = method)) +
  geom_hline(yintercept = .05, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Type I Error", limits = c(.025, .105), breaks = seq(0.03, .09, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = power, group = method)) +
  geom_hline(yintercept = .05, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Type I Error", limits = c(.025, .105), breaks = seq(0.03, .09, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = power, group = method)) +
  geom_hline(yintercept = .05, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Type I Error", limits = c(.025, .105), breaks = seq(0.03, .09, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

```


#### Power (True Small Effect)

```{r}
#| label: plotOverallPowerSmall
#| echo: false
#| eval: true

plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = power, group = method)) +
  geom_hline(yintercept = .05, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Power", limits = c(.1, .6), breaks = seq(.1, .6, .1)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = power, group = method)) +
  geom_hline(yintercept = .05, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Power", limits = c(.1, .6), breaks = seq(.1, .6, .1)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = power, group = method)) +
  geom_hline(yintercept = .05, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Power", limits = c(.1, .6), breaks = seq(.1, .6, .1)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

```


#### Power (True Large Effect)

```{r}
#| label: plotOverallPowerMedium
#| echo: false
#| eval: true

plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = power, group = method)) +
  geom_hline(yintercept = .05, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Power", limits = c(.35, 1), breaks = seq(0.4, 1, .2)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = power, group = method)) +
  geom_hline(yintercept = .05, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Power", limits = c(.35, 1), breaks = seq(0.4, 1, .2)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = power, group = method)) +
  geom_hline(yintercept = .05, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Power", limits = c(.35, 1), breaks = seq(0.4, 1, .2)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

```


### Figures: Bias

```{r}
#| label: plotOptions2
#| echo: false
#| eval: true

# Filter plot_data to exclude methods without CI
plot_dat <- filter(plot_dat, !(method %in% c("cmh test")))

# Gather methods and organize in clusters
breaks <- c(
  "Ge (orig)", "Liu (HC3)", "Ye", "Firth (FLIC)", "g-comp bootstrap", "Score Test",
  "MH (mGR)", "MH (Sato)", "suissa-shuster"
)

# Set labels
labels <- c(
  "G-Comp.: Ge",
  "G-Comp.: Liu",
  "G-Comp.: Ye",
  "G-Comp.: Firth",
  "G-Comp.: Bootstrap",
  "G-Comp.: Score Test",
  "Mantel-Haenszel RD: modified GR",
  "Mantel-Haenszel RD: Sato",
  "Suissa-Shuster Test"
)

# Set colors: unique per method, grouped by clusters
values <- c(
  "dodgerblue4", "lightskyblue1", "deepskyblue", "dodgerblue", "#5DADE2", "blue",  # G-Comp (blue family)
  "#E69F00", "#F0E442",                                              # MH (orange/yellow)
  "#009E73"                                                          # CMH (green)
)

# Linetypes: Same within cluster, different across clusters
line_types <- c(
  rep("solid", 6),      # G-Comp
  rep("dashed", 2),     # MH
  "dotted"              # Suissa-Shuster
)

# Shapes: Unique within clusters
shapes <- c(19, 17, 15, 3, 8, 18, 19, 17, 19)
```

#### True Null Effect

**Caution**: Please note the narrow range on the y-axis. The tight bounds are intentional to highlight minor differences between methods, should they exist.

```{r}
#| label: plotBiasNull
#| echo: false
#| eval: true
#| width: 9

r <- plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = rmse, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("RMSE", limits = c(0, .16) , breaks = seq(0, .15, .05)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

b <- plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = bias_mean, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Bias", limits = c(-.0015, .0015)
                     , breaks = seq(-.0015, .0015, .0015)
                     ) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

ggpubr::ggarrange(b, r, nrow = 1, common.legend = TRUE, legend = "bottom", align = "h")


r <- plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = rmse, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("RMSE", limits = c(0, .16) , breaks = seq(0, .15, .05)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

b <- plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = bias_mean, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Bias", limits = c(-.0015, .0015)
                     , breaks = seq(-.0015, .0015, .0015)
  ) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

ggpubr::ggarrange(b, r, nrow = 1, common.legend = TRUE, legend = "bottom", align = "h")


r <- plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = rmse, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("RMSE", limits = c(0, .16) , breaks = seq(0, .15, .05)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

b <- plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = bias_mean, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Bias", limits = c(-.0015, .0015)
                     , breaks = seq(-.0015, .0015, .0015)
  ) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

ggpubr::ggarrange(b, r, nrow = 1, common.legend = TRUE, legend = "bottom", align = "h")

```


#### True Medium Effect

**Caution**: Please note the change in the y-axis range.

```{r}
#| label: plotBiasMedium
#| echo: false
#| eval: true

r <- plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = rmse, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("RMSE", limits = c(0, .2) , breaks = seq(0, .2, .05)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

b <- plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = bias_mean, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Bias", limits = c(-.022, .005) , breaks = seq(-.02, .005, .005)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

ggpubr::ggarrange(b, r, nrow = 1, common.legend = TRUE, legend = "bottom", align = "h")


r <- plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = rmse, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("RMSE", limits = c(0, .2) , breaks = seq(0, .2, .05)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

b <- plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = bias_mean, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Bias", limits = c(-.022, .005) , breaks = seq(-.02, .005, .005)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

ggpubr::ggarrange(b, r, nrow = 1, common.legend = TRUE, legend = "bottom")


r <- plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = rmse, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("RMSE", limits = c(0, .2) , breaks = seq(0, .2, .05)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

b <- plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = bias_mean, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Bias", limits = c(-.022, .005) , breaks = seq(-.02, .005, .005)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

ggpubr::ggarrange(b, r, nrow = 1, common.legend = TRUE, legend = "bottom")
```

#### True Large Effect

**Caution**: Please note the change in the y-axis range.

```{r}
#| label: plotBiasLarge
#| echo: false
#| eval: true

r <- plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = rmse, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("RMSE", limits = c(0, .2) , breaks = seq(0, .2, .05)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

b <- plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = bias_mean, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Bias", limits = c(-.04, .005) , breaks = seq(-.04, .005, .005)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

ggpubr::ggarrange(b, r, nrow = 1, common.legend = TRUE, legend = "bottom")


r <- plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = rmse, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("RMSE", limits = c(0, .2) , breaks = seq(0, .2, .05)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

b <- plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = bias_mean, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Bias", limits = c(-.04, .005) , breaks = seq(-.04, .005, .005)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

ggpubr::ggarrange(b, r, nrow = 1, common.legend = TRUE, legend = "bottom")


r <- plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = rmse, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("RMSE", limits = c(0, .2) , breaks = seq(0, .2, .05)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

b <- plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = bias_mean, group = method)) +
  geom_hline(yintercept = 0, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Bias", limits = c(-.04, .005) , breaks = seq(-.04, .005, .005)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

ggpubr::ggarrange(b, r, nrow = 1, common.legend = TRUE, legend = "bottom")
```


### Figures: Coverage

```{r}
#| label: plotOptions3
#| echo: false
#| eval: true

# Filter plot_data to exclude methods without CI
plot_dat <- filter(plot_dat, !(method %in% c("suissa-shuster", "cmh test")))

# Gather methods and organize in clusters
breaks <- c(
  "Ge (orig)", "Liu (HC3)", "Ye", "Firth (FLIC)", "g-comp bootstrap", "Score Test",
  "MH (mGR)", "MH (Sato)"
)

# Set labels
labels <- c(
  "G-Comp.: Ge",
  "G-Comp.: Liu",
  "G-Comp.: Ye",
  "G-Comp.: Firth",
  "G-Comp.: Bootstrap",
  "G-Comp.: Score Test",
  "Mantel-Haenszel RD: modified GR",
  "Mantel-Haenszel RD: Sato"
)

# Set colors: unique per method, grouped by clusters
values <- c(
  "dodgerblue4", "lightskyblue1", "deepskyblue", "dodgerblue", "#5DADE2", "blue",  # G-Comp (blue family)
  "#E69F00", "#F0E442"                                              # MH (orange/yellow)
)

# Linetypes: Same within cluster, different across clusters
line_types <- c(
  rep("solid", 6),      # G-Comp
  rep("dashed", 2)     # MH
)

# Shapes: Unique within clusters
shapes <- c(19, 17, 15, 3, 8, 18, 19, 17)
```


#### True Null Effect

```{r}
#| label: plotCoverageNull
#| echo: false
#| eval: true

plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = coverage, group = method)) +
  geom_hline(yintercept = .95, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Coverage", limits = c(.9, .98), breaks = seq(0.9, .98, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = coverage, group = method)) +
  geom_hline(yintercept = .95, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Coverage", limits = c(.9, .98), breaks = seq(0.9, .98, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = coverage, group = method)) +
  geom_hline(yintercept = .95, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Coverage", limits = c(.9, .98), breaks = seq(0.9, .98, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

```

#### True Medium Effect

```{r}
#| label: plotCoverageMedium
#| echo: false
#| eval: true

plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = coverage, group = method)) +
  geom_hline(yintercept = .95, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Coverage", limits = c(.9, .98), breaks = seq(0.9, .98, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = coverage, group = method)) +
  geom_hline(yintercept = .95, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Coverage", limits = c(.9, .98), breaks = seq(0.9, .98, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0.15, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = coverage, group = method)) +
  geom_hline(yintercept = .95, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Coverage", limits = c(.9, .98), breaks = seq(0.9, .98, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.15; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

```

#### True Large Effect

```{r}
#| label: plotCoverageLarge
#| echo: false
#| eval: true

plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: None") %>%
  ggplot(aes(x = n, y = coverage, group = method)) +
  geom_hline(yintercept = .95, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Coverage", limits = c(.9, .98), breaks = seq(0.9, .98, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: None") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: Medium") %>%
  ggplot(aes(x = n, y = coverage, group = method)) +
  geom_hline(yintercept = .95, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Coverage", limits = c(.9, .98), breaks = seq(0.9, .98, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: Medium") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

plot_dat %>%
  filter(diff == 0.30, cov_prog_val == "Prog. Val: High") %>%
  ggplot(aes(x = n, y = coverage, group = method)) +
  geom_hline(yintercept = .95, lwd = .75, linetype = "dashed") +
  geom_line(aes(color = method, linetype = method), linewidth = .8,
            position = position_dodge(width = 3)) +
  geom_point(aes(color = method, shape = method), size = 3,
             position = position_dodge(width = 3)) +
  scale_y_continuous("Coverage", limits = c(.9, .98), breaks = seq(0.9, .98, .02)) +
  scale_x_continuous(limits = c(20, 160), breaks = seq(30, 150, 30)) +
  scale_color_manual("Method", breaks = breaks, values = values, labels = labels) +
  scale_linetype_manual("Method", breaks = breaks, values = line_types, labels = labels) +
  scale_shape_manual("Method", breaks = breaks, values = shapes, labels = labels) +
  labs(x = "Total Sample Size",
       title = "Risk Difference = 0.30; Prog. Val. Covariate: High") +
  theme_bw() +
  theme(
    text = element_text(size = 10, color = "black"),
    legend.position = "right",
    legend.key.width = unit(1.5, "lines"),
    axis.text = element_text(size = 10, color = "black")
  )

```
